---
title: "HW2_315"
author: "Justin Song - jjs5874 - Scott - https://github.com/jujiwoo/HW2_315/blob/main/HW2%20315.Rmd"
date: "2024-01-23"
output:
  pdf_document: default
  html_document: default
---

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(fig.height=5, fig.width=10, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=100))
```


```{r, echo=FALSE, results='hide', message=FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(knitr)
profs = read.csv('profs.csv')
bike = read.csv('bikeshare.csv')
metro = read.csv('capmetro_UT.csv')
billboard = read.csv('billboard.csv')
```
**Problem 1: Beauty, or not, in the classroom**


Part A. 

```{r, echo=FALSE, results='hide', message=FALSE}
ggplot(profs, aes(eval)) +  geom_histogram(binwidth = 0.5, col='black', fill='maroon') + xlab('Course Evaluation Scores') + ylab('Frequency') + ggtitle('Distribution of Evaluation Scores for Professors') + theme_classic()
```

The distribution of course evaluation scores for professors is skewed left and has a peak around 4. This indicates that most professors received an average evaluation scores in the higher range. This histogram essentially shows that UT students generally hold a highly positive attractive perception towards their professors.

Part B. 

```{r, echo=FALSE, results='hide', message=FALSE}
ggplot(profs, aes(eval)) + 
  geom_boxplot(col='black', fill='maroon') + 
  xlab('Course Evaluation Scores') + 
  ylab('Frequency') + 
  ggtitle('Distribution of Evaluation Scores for Professors') + 
  facet_wrap(~native, labeller = labeller(native = c("no" = "Not a Native English Speaker", "yes" = "Native English Speaker"))) + 
  theme_classic()

```
This boxplot shows the distribution of evaluation scores for native and non-native professors. The distributions have similar shapes but native professors tend to have slightly higher median evaluation scores.


Part C. 

```{r, echo=FALSE, results='hide', message=FALSE}
ggplot(profs, aes(eval)) +  geom_histogram(binwidth = 0.5, col='black', fill='maroon') + xlab('Course Evaluation Scores') + ylab('Frequency') + ggtitle('Distribution of Evaluation Scores for Professors') + facet_wrap(~gender, nrow=2, labeller = labeller(gender = c(female = "Female", male = "Male"))) + theme_classic()
```
These histograms of evaluation scores grouped by professor gender show that both male and female professors have left-skewed distributions peaked near 4. The distributions appear fairly similar for both genders, however, male professors have a noticable higher course evaluation average. 


Part D.

```{r, echo=FALSE, results='hide', message=FALSE}
ggplot(profs, aes(eval, beauty)) +  geom_point(col='maroon') + xlab('Course Evaluation Scores') + ylab('Beauty Score') + ggtitle('Relationship of Evaluation Scores for Professors and Beauty Score') + theme_classic()
```
This scatterplot shows there is a positive, but unclear linear relationship between professors' course evaluation scores and their beauty scores. The points are widely scattered, indicating beauty score does not strongly predict evaluation score.

**Problem 2: bike sharing**



 Plot A:                                                                                                                                                                             
```{r, echo=FALSE, results='hide', message=FALSE}
bike %>%
  group_by(hr) %>%
  summarize(avg_hr_bike = mean(total)) %>%
  ggplot(aes(x=hr, y=avg_hr_bike)) + geom_line() + labs(title="Change of Hourly Bike Rentals Throughout the Day", x = "Hour", y = "Average Bike Rentals")
```
This line graph shows the average hourly bike rentals across all hours of the day. The x-axis displays the hour of the day from 0 to 23, and the y-axis shows the average number of total bike rentals for each hour. We can see that rentals peak around 8-9am and 5-6pm, corresponding to rush hour commute times. Overall, the plot illustrates strong commuting patterns in bike usage.


Plot B:

```{r, echo=FALSE, results='hide', message=FALSE, message=FALSE}
bike %>%
  group_by(hr, workingday) %>%
  summarize(avg_hr_bike = mean(total)) %>%
  ggplot(aes(x = hr, y = avg_hr_bike)) + 
  geom_line() + 
  facet_wrap(~workingday, labeller = labeller(workingday = c("0" = "Not a Working Day", "1" = "Working Day"))) + labs(title="Change of Hourly Bike Rentals Throughout the Day", x = "Hour", y = "Average Bike Rentals")

```

This faceted line graph displays the average bike rentals by hour, separated into working days vs non-working days. The x-axis shows the hour of day from 0 to 23, the y-axis shows the average number of total bike rentals, and the panels split the data based on working day (1) or not (0). We can see pronounced morning and evening rush hours on working days, while weekends show steadier rental patterns throughout the day. This illustrates how commuting heavily influences bike rental demand during the work week.


Plot C: 

```{r, echo=FALSE, results='hide', message=FALSE}
bike %>%
  filter(hr == 9) %>%
  group_by(weathersit, workingday) %>%
  summarize(avg_ridership = mean(total)) %>%

  ggplot(aes(x = weathersit, y = avg_ridership)) +
  geom_col() +
  facet_wrap(~workingday, labeller = labeller(workingday = c("0" = "Not a Working Day", "1" = "Working Day"))) + 
  labs(
    title = "Average 9 AM Ridership by Weather and Working Day",
    x = "Weather Situation",
    y = "Average Ridership"
  )
```

This faceted bar chart focuses on ridership patterns at 9am based on weather and working day. The x-axis shows weather situation (clear, mist, light precipitation, heavy precipitation), the y-axis shows the average number of rentals, and the panels separate working days and non-working days. We see higher 9am demand on clear days, especially on working days. The plot indicates weather and work commuting both influence bike rental patterns in the morning.


**Problem 3 - Capital Metro UT Ridership**

```{r, echo=FALSE, results='hide', message=FALSE}
# Recode the categorical variables in sensible, rather than alphabetical, order
metro = mutate(metro,
day_of_week = factor(day_of_week,
levels=c("Mon", "Tue", "Wed","Thu", "Fri", "Sat", "Sun")),
month = factor(month,
levels=c("Sep", "Oct","Nov")))
```

Plot 1: 

```{r, echo=FALSE, results='hide', message=FALSE}
metro %>%
  group_by(hour_of_day, day_of_week, month) %>%
  summarize(avg_boardings = mean(boarding)) %>%
  ggplot(aes(x=hour_of_day, y=avg_boardings, color=month)) + geom_line() + facet_wrap(~day_of_week) + labs(title = "Average Boardings by Hour of the Day, Day of Week, and Month", x="Hour of Day", y = "Average Boardings")
```

Average Metro Boardings by Time, Day, and Month. This faceted line chart displays average hourly ridership across days of the week and months. The hour of peak boardings is fairly consistent, typically 8-9am on weekdays and 11am-12pm on weekends. However, total daily ridership varies. Mondays in September have lower averages, likely due to the start of the semester. Dips in ridership also occur on Wednesdays, Thursdays, and Fridays in November, probably due to the Thanksgiving holiday break. While the hourly boarding pattern is steady, factors like academic schedules influence total ridership over the course of the semester.


Plot 2:

```{r, echo=FALSE, results='hide', message=FALSE}
ggplot(metro, aes(x = temperature, y = boarding, color = factor(weekend)), size = 0.5) +
  geom_point() + 
  facet_wrap(~hour_of_day) +
  labs(
    title = "UT Student Metro Ridership by Temperature and Time",
    x = "Temperature (F)",
    y = "Boardings",
    color = "Weekday vs Weekend")
```

This faceted scatterplot shows the relationship between temperature and student ridership on the UT metro system. Each panel shows a different hour of the day, with points colored by weekday (blue) or weekend (red). When holding hour of day and weekend status constant, there does not appear to be a strong correlation between temperature and ridership. For example, focusing on the 7am weekday panel, we see similar ridership across the range of temperatures.

**Problem 4: Wrangling the Billboard Top 100**

Part A: 

```{r, echo=FALSE, message =FALSE}
top10 = billboard %>% 
  group_by(performer, song) %>%
  summarise(weeks = sum(week, na.rm = TRUE)) %>%
  arrange(desc(weeks)) %>%
  head(10)

kable(top10, col.names = c("Performer", "Song", "Count"))
```
This table displays the top 10 songs that have spent the most cumulative weeks on the Billboard Hot 100 chart since 1958.


Part B: 

```{r, echo=FALSE, message=FALSE}
billboard_no_outliers = billboard %>% 
  filter(year != 1958, year != 2021)

yearly_songs = billboard_no_outliers %>%
  group_by(year) %>%
  summarise(unique_songs = length(unique(song)))

ggplot(yearly_songs, aes(x = year, y = unique_songs)) +
  geom_line() +
  labs(title = "Musical Diversity of Billboard Top 100 Over Time",
       x = "Year",
       y = "Number of Unique Songs")
```
This figure shows the number of unique songs appearing on the Billboard Top 100 chart each year from 1959 to 2020. There is an overall upward trend, indicating increasing musical diversity over time, with some particularly steep increases in diversity in the 1960s, 1990s, and 2010s.



Part C: 

```{r, echo=FALSE, results='hide', message=FALSE}
ten_week_songs = billboard %>% 
  filter(weeks_on_chart >= 10)

ten_week_hits = ten_week_songs %>%
  group_by(performer) %>%
  summarise(ten_week_hits = length(unique((song))))

ten_week_organized = ten_week_hits %>%
    arrange(desc(ten_week_hits))

top_19 = ten_week_organized %>% top_n(19, ten_week_hits)


ggplot(top_19, aes(x = reorder(performer, ten_week_hits), y = ten_week_hits)) +
  geom_col() + theme(axis.text.x = element_text(angle = 90)) +
  labs(
    title = "Artists with 30+ Ten-Week Hits on Billboard Top 100",
    x = "Artist",
    y = "Number of Ten-Week Hits")
```
This bar chart shows the 19 artists with the most 'ten-week hits' (songs that charted on the Billboard Top 100 for at least 10 weeks) since 1958. Elton John tops the list with 39 ten-week hits.






